---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## radomd

R wrapper for Python package `Pyadomd` https://github.com/S-C-O-U-T/Pyadomd. An Rthonic on top of the pythonic approach to query SSAS data models. 

## Installation and set up

1. Install `ADOMD`client library from https://learn.microsoft.com/en-us/analysis-services/client-libraries?view=asallproducts-allversions

2. Install `Reticulate` R library

```{r, eval=FALSE}
install.packages('reticulate')
```

3. Install `Pyadomd` Python library to be used by `Reticulate` in R.

```{r, eval=FALSE}
library(reticulate)
repl_python(input = "!pip install pyadomd")
```

4. Install `Radomd` R library:

```{r, eval=FALSE}
devtools::install_github("JacekPardyak/radomd")
```

5. Set up environmental variables:

```{r}
Sys.setenv(adomd_path = "/Program Files/Microsoft.NET/ADOMD.NET/160",
           conn_str = "Data Source=localhost; Provider=MSOLAP; initial catalog=Analysis Services Tutorial")
```

6. Check if everything works fine:

```{r, eval=FALSE}
library(radomd)
Query() %>% execute()
# Open
```


## Query SSAS Tabular model with R

1. You can use creator:

```{r, eval=FALSE}
library(radomd)
Sys.setenv(adomd_path = "/Program Files/Microsoft.NET/ADOMD.NET/160",
           conn_str = "Data Source=localhost; Provider=MSOLAP; initial catalog=Analysis Services Tutorial")
Query() %>% cube("[Adventure Works]") %>%
  rows(c("[Product].[Product Line].[Product Line].MEMBERS")) %>% 
  columns(c("[Measures].[Internet Sales Count]", "[Measures].[Internet Sales-Sales Amount]")) %>%
  slicers(c("[Sales Territory].[Sales Territory Country].[Australia]")) %>% execute()
```

2. Or you can use MDX to get the same result:

```{r, eval=FALSE}
library(radomd)
Query() %>% mdx("SELECT {[Measures].[Internet Sales Count],
[Measures].[InternetSales-Sales Amount]} ON COLUMNS, 
{[Product].[Product Line].[Product Line].MEMBERS} ON ROWS
FROM [Analysis Services Tutorial] 
WHERE [Sales Territory].[Sales Territory Country].[Australia]") %>% execute()
```

